---
title: "Figure 3a, S1a, S2a"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
rm(list = ls())

library(tidyverse)
library(phyloseq)
library(ggpubr)
library(pander)
panderOptions('table.caption.prefix', NULL)
panderOptions('table.continues', NULL)
panderOptions('table.emphasize.rownames', FALSE)

gg_color_hue=function(n){
  hues = seq(15, 375, length = n + 1)
  hcl(h = hues, l = 65, c = 100)[1:n]
}

source("ancom_bc_v1.0.R")
source("sim_data_poi_gam_two_grp.R")
```

# 1. Generate Simulation Data

## 1.1 Large variable sampling fractions

Unbalanced microbial loads and balanced library sizes

```{r, message=FALSE, warning=FALSE, comment=NA}
abn.tab.gen1=function(n.taxa, n.samp.grp1, n.samp.grp2, prop.diff, abn.seed, obs.seed, 
                      struc.zero.prop, out.zero.prop, library.size){
  # Total number of samples
  n.samp=n.samp.grp1+n.samp.grp2
  
  set.seed(abn.seed) # This seed is used to control whether you would like to have the same population
  low.prop=0.6 # Proportion of low abundance 
  med.prop=0.3 # Proportion of medium abundance
  hi.prop=0.1  # Proportion of high abundance
  # Indices for taxa abundance 
  index=sample(c(1, 2, 3), n.taxa, replace = T, prob = c(low.prop, med.prop, hi.prop)) 
  
  # Poisson parameters
  lambda=rep(NA, n.taxa)
  lambda[which(index==1)]=rgamma(length(which(index==1)), shape=50, rate=1)
  lambda[which(index==2)]=rgamma(length(which(index==2)), shape=200, rate=1)
  lambda[which(index==3)]=rgamma(length(which(index==3)), shape=10000, rate=1)
  
  # Which taxa are differentially abundant
  diff.ind=rep(0, n.taxa)
  # Group1 is higher than group2
  diff1.ind=sample(c(1:n.taxa), floor(n.taxa*prop.diff), replace=FALSE)
  diff.ind[diff1.ind]=1
  # Group2 is higher than group1
  wt=runif(1, 0, 1)
  diff2.ind=sample(diff1.ind, wt*length(diff1.ind), replace=FALSE)
  diff.ind[diff2.ind]=2
  # Structural zeros
  diff3.ind=sample(which(diff.ind==0), struc.zero.prop*length(which(diff.ind==0)), replace = FALSE)
  diff.ind[diff3.ind]=-1
  
  # Effect size
  effect.size=rep(1, n.taxa)
  effect.size[diff1.ind]=runif(length(diff1.ind), 1, 10)
  effect.size[diff2.ind]=runif(length(diff2.ind), 0.1, 1)
  effect.size[diff3.ind]=0
  names(effect.size)=paste0("taxon", seq(n.taxa))
  
  # Mean absolute abundance in the ecosystem
  temp.grp1=round(lambda*effect.size)
  temp.grp2=round(lambda)
  for (i in which(effect.size!=1)) {
    if(temp.grp1[i]==temp.grp2[i]) temp.grp1[i]=temp.grp1[i]+1
  }
  temp.dat=data.frame(temp.grp1, temp.grp2, effect.size)
  rownames(temp.dat)=paste0("taxon", seq(n.taxa))
  
  # Absolute abundance in the ecosystem
  abn.mat=matrix(0, ncol=n.samp, nrow=n.taxa)
  for(i in 1:n.taxa){
    abn.mat[i, ]=c(rpois(n.samp.grp1, temp.grp1[i]), rpois(n.samp.grp2, temp.grp2[i]))
  }
  # Outlier zeros
  out.ind=rep(0, n.taxa); out.ind[sample(seq(n.taxa), out.zero.prop*n.taxa, replace = F)]=1
  names(out.ind)=paste0("taxon", seq(n.taxa))
  abn.mat[which(out.ind==1), sample(seq(n.samp), out.zero.prop*n.samp, replace = F)]=0
  
  abn.total=colSums(abn.mat)
  names(abn.total)=paste0("sub", seq(n.samp))
  
  # Library size
  if(library.size=="large"){
    depth=1/runif(n.samp, 5, 10)
    obs.total=round(max(abn.total)*depth)
  }else if(library.size=="small"){
    depth=1/sample(c(runif(n.samp, 10, 50), runif(n.samp, 100, 500)), n.samp, replace = T)
    obs.total=round(max(abn.total)*depth)
  }
  names(obs.total)=paste0("sub", seq(n.samp))
  
  # Absolute abundance in a sample
  set.seed(obs.seed)
  obs.list=lapply(1:n.samp, function(i) 
    phyloseq:::rarefaction_subsample(x=abn.mat[, i], sample.size=obs.total[i]))
  obs.mat=Reduce('cbind', obs.list)
  
  # Prepare outputs
  abn.dat=data.frame(abn.mat, row.names = NULL)
  rownames(abn.dat)=paste0("taxon", seq(n.taxa))
  colnames(abn.dat)=paste0("sub", seq(n.samp))
  
  obs.dat=data.frame(obs.mat, row.names = NULL)
  rownames(obs.dat)=paste0("taxon", seq(n.taxa))
  colnames(obs.dat)=paste0("sub", seq(n.samp))
  
  grp.ind=c(rep(1, n.samp.grp1), rep(2, n.samp.grp2))
  names(grp.ind)=paste0("sub", seq(n.samp))
  
  names(diff.ind)=paste0("taxon", seq(n.taxa))
  
  c.mult=obs.total/abn.total
  names(c.mult)=paste0("sub", seq(n.samp))
  
  test.data=list(temp.dat, abn.dat, obs.dat, effect.size, grp.ind, 
                 diff.ind, out.ind, c.mult, abn.total, obs.total)
  names(test.data)=c("mean.eco.abn", "eco.abn", "obs.abn", "effect.size", "grp", 
                     "diff.taxa", "outlier", "mult", "micro.load", "lib.size")
  return(test.data)
}
```

## 1.2 Moderate variable sampling fractions

Unbalanced microbial loads and unbalanced library sizes

```{r, message=FALSE, warning=FALSE, comment=NA}
abn.tab.gen2=function(n.taxa, n.samp.grp1, n.samp.grp2, prop.diff, abn.seed, obs.seed,
                      struc.zero.prop, out.zero.prop, library.size){
  # Total number of samples
  n.samp=n.samp.grp1+n.samp.grp2
  
  set.seed(abn.seed) # This seed is used to control whether you would like to have the same population
  low.prop=0.6 # Proportion of low abundance 
  med.prop=0.3 # Proportion of medium abundance
  hi.prop=0.1  # Proportion of high abundance
  # Indices for taxa abundance 
  index=sample(c(1, 2, 3), n.taxa, replace = T, prob = c(low.prop, med.prop, hi.prop)) 
  
  # Poisson parameters
  lambda=rep(NA, n.taxa)
  lambda[which(index==1)]=rgamma(length(which(index==1)), shape=50, rate=1)
  lambda[which(index==2)]=rgamma(length(which(index==2)), shape=200, rate=1)
  lambda[which(index==3)]=rgamma(length(which(index==3)), shape=10000, rate=1)
  
  # Which taxa are differentially abundant
  diff.ind=rep(0, n.taxa)
  # Group1 is higher than group2
  diff1.ind=sample(c(1:n.taxa), floor(n.taxa*prop.diff), replace=FALSE)
  diff.ind[diff1.ind]=1
  # Group2 is higher than group1
  wt=runif(1, 0, 1)
  diff2.ind=sample(diff1.ind, wt*length(diff1.ind), replace=FALSE)
  diff.ind[diff2.ind]=2
  # Structural zeros
  diff3.ind=sample(which(diff.ind==0), struc.zero.prop*length(which(diff.ind==0)), replace = FALSE)
  diff.ind[diff3.ind]=-1
  
  # Effect size
  effect.size=rep(1, n.taxa)
  effect.size[diff1.ind]=runif(length(diff1.ind), 1, 10)
  effect.size[diff2.ind]=runif(length(diff2.ind), 0.1, 1)
  effect.size[diff3.ind]=0
  names(effect.size)=paste0("taxon", seq(n.taxa))
  
  # Mean absolute abundance in the ecosystem
  temp.grp1=round(lambda*effect.size)
  temp.grp2=round(lambda)
  for (i in which(effect.size!=1)) {
    if(temp.grp1[i]==temp.grp2[i]) temp.grp1[i]=temp.grp1[i]+1
  }
  temp.dat=data.frame(temp.grp1, temp.grp2, effect.size)
  rownames(temp.dat)=paste0("taxon", seq(n.taxa))
  
  # Absolute abundance in the ecosystem
  abn.mat=matrix(0, ncol=n.samp, nrow=n.taxa)
  for(i in 1:n.taxa){
    abn.mat[i, ]=c(rpois(n.samp.grp1, temp.grp1[i]), rpois(n.samp.grp2, temp.grp2[i]))
  }
  # Outlier zeros
  out.ind=rep(0, n.taxa); out.ind[sample(seq(n.taxa), out.zero.prop*n.taxa, replace = F)]=1
  names(out.ind)=paste0("taxon", seq(n.taxa))
  abn.mat[which(out.ind==1), sample(seq(n.samp), out.zero.prop*n.samp, replace = F)]=0
  
  abn.total=colSums(abn.mat)
  names(abn.total)=paste0("sub", seq(n.samp))
  
  # Library size
  if(library.size=="large"){
    depth=1/runif(n.samp, 5, 10)
    obs.total=round(max(abn.total)*depth)
  }else if(library.size=="small"){
    depth=1/sample(c(runif(n.samp, 10, 50), runif(n.samp, 100, 500)), n.samp, replace = T)
    obs.total=round(abn.total*depth)
  }
  names(obs.total)=paste0("sub", seq(n.samp))
  
  # Absolute abundance in a sample
  set.seed(obs.seed)
  obs.list=lapply(1:n.samp, function(i) 
    phyloseq:::rarefaction_subsample(x=abn.mat[, i], sample.size=obs.total[i]))
  obs.mat=Reduce('cbind', obs.list)
  
  # Prepare outputs
  abn.dat=data.frame(abn.mat, row.names = NULL)
  rownames(abn.dat)=paste0("taxon", seq(n.taxa))
  colnames(abn.dat)=paste0("sub", seq(n.samp))
  
  obs.dat=data.frame(obs.mat, row.names = NULL)
  rownames(obs.dat)=paste0("taxon", seq(n.taxa))
  colnames(obs.dat)=paste0("sub", seq(n.samp))
  
  grp.ind=c(rep(1, n.samp.grp1), rep(2, n.samp.grp2))
  names(grp.ind)=paste0("sub", seq(n.samp))
  
  names(diff.ind)=paste0("taxon", seq(n.taxa))
  
  c.mult=obs.total/abn.total
  names(c.mult)=paste0("sub", seq(n.samp))
  
  test.data=list(temp.dat, abn.dat, obs.dat, effect.size, grp.ind, 
                 diff.ind, out.ind, c.mult, abn.total, obs.total)
  names(test.data)=c("mean.eco.abn", "eco.abn", "obs.abn", "effect.size", "grp", 
                     "diff.taxa", "outlier", "mult", "micro.load", "lib.size")
  return(test.data)
}
```

## 1.3 Small variable sampling fractions

Balanced microbial loads and balanced library sizes

```{r, message=FALSE, warning=FALSE, comment=NA}
abn.tab.gen3=function(n.taxa, n.samp.grp1, n.samp.grp2, prop.diff, abn.seed, obs.seed,
                      struc.zero.prop, out.zero.prop, library.size){
  # Total number of samples
  n.samp=n.samp.grp1+n.samp.grp2
  
  set.seed(abn.seed) # This seed is used to control whether you would like to have the same population
  low.prop=0.6 # Proportion of low abundance 
  med.prop=0.3 # Proportion of medium abundance
  hi.prop=0.1  # Proportion of high abundance
  # Indices for taxa abundance 
  index=sample(c(1, 2, 3), n.taxa, replace = T, prob = c(low.prop, med.prop, hi.prop)) 
  
  # Poisson parameters
  lambda=rep(NA, n.taxa)
  lambda[which(index==1)]=rgamma(length(which(index==1)), shape=50, rate=1)
  lambda[which(index==2)]=rgamma(length(which(index==2)), shape=200, rate=1)
  lambda[which(index==3)]=rgamma(length(which(index==3)), shape=10000, rate=1)
  
  # Which taxa are differentially abundant
  diff.ind=rep(0, n.taxa)
  # Differentially abundant taxa
  diff.pos=sample(c(1:n.taxa), floor(n.taxa*prop.diff), replace=FALSE)
  diff.ind[diff.pos]=1
  # Structural zeros
  szero.pos=sample(which(diff.ind==0), struc.zero.prop*length(which(diff.ind==0)), replace = FALSE)
  diff.ind[szero.pos]=-1
  
  # Effect size
  effect.size=rep(1, n.taxa)
  effect.size[diff.pos]=runif(length(diff.pos), 1, 10)
  effect.size[szero.pos]=0
  names(effect.size)=paste0("taxon", seq(n.taxa))
  
  # Mean absolute abundance in the ecosystem
  temp.grp1=round(lambda); temp.grp2=round(lambda)
  for (i in diff.pos) {
    rand.ind=sample(c(1, 2), 1)
    if (rand.ind==1){
      temp.grp1[i]=temp.grp1[i]*effect.size[i]
    }else{
      temp.grp2[i]=temp.grp2[i]*effect.size[i]
    }
  }
  for (i in szero.pos) {
    rand.ind=sample(c(1, 2), 1)
    if (rand.ind==1){
      temp.grp1[i]=temp.grp1[i]*effect.size[i]
    }else{
      temp.grp2[i]=temp.grp2[i]*effect.size[i]
    }
  }
  temp.dat=data.frame(temp.grp1, temp.grp2, effect.size)
  rownames(temp.dat)=paste0("taxon", seq(n.taxa))
  
  # Absolute abundance in the ecosystem
  abn.mat=matrix(0, ncol=n.samp, nrow=n.taxa)
  for(i in 1:n.taxa){
    abn.mat[i, ]=c(rpois(n.samp.grp1, temp.grp1[i]), rpois(n.samp.grp2, temp.grp2[i]))
  }
  # Outlier zeros
  out.ind=rep(0, n.taxa); out.ind[sample(seq(n.taxa), out.zero.prop*n.taxa, replace = F)]=1
  names(out.ind)=paste0("taxon", seq(n.taxa))
  abn.mat[which(out.ind==1), sample(seq(n.samp), out.zero.prop*n.samp, replace = F)]=0
  
  abn.total=colSums(abn.mat)
  names(abn.total)=paste0("sub", seq(n.samp))
  
  # Library size
  if(library.size=="large"){
    depth=1/runif(n.samp, 5, 10)
    obs.total=round(max(abn.total)*depth)
  }else if(library.size=="small"){
    depth=1/sample(c(runif(n.samp, 10, 50), runif(n.samp, 100, 500)), n.samp, replace = T)
    obs.total=round(max(abn.total)*depth)
  }
  names(obs.total)=paste0("sub", seq(n.samp))
  
  # Absolute abundance in a sample
  set.seed(obs.seed)
  obs.list=lapply(1:n.samp, function(i) 
    phyloseq:::rarefaction_subsample(x=abn.mat[, i], sample.size=obs.total[i]))
  obs.mat=Reduce('cbind', obs.list)
  
  # Prepare outputs
  abn.dat=data.frame(abn.mat, row.names = NULL)
  rownames(abn.dat)=paste0("taxon", seq(n.taxa))
  colnames(abn.dat)=paste0("sub", seq(n.samp))
  
  obs.dat=data.frame(obs.mat, row.names = NULL)
  rownames(obs.dat)=paste0("taxon", seq(n.taxa))
  colnames(obs.dat)=paste0("sub", seq(n.samp))
  
  grp.ind=c(rep(1, n.samp.grp1), rep(2, n.samp.grp2))
  names(grp.ind)=paste0("sub", seq(n.samp))
  
  names(diff.ind)=paste0("taxon", seq(n.taxa))
  
  c.mult=obs.total/abn.total
  names(c.mult)=paste0("sub", seq(n.samp))
  
  test.data=list(temp.dat, abn.dat, obs.dat, effect.size, grp.ind, 
                 diff.ind, out.ind, c.mult, abn.total, obs.total)
  names(test.data)=c("mean.eco.abn", "eco.abn", "obs.abn", "effect.size", "grp", 
                     "diff.taxa", "outlier", "mult", "micro.load", "lib.size")
  return(test.data)
}
```

# 2. Normalization Efficacy with the Presence of Large Variable Sampling Fractions

## 2.1 Run a single simulation dataset

```{r, message=FALSE, warning=FALSE, comment=NA, eval=FALSE}
n.taxa=500; n.samp.grp1=30; n.samp.grp2=30; low.abn=50; med.abn=200; high.abn=10000
prop.diff=0.25; abn.seed=123; obs.seed=124; struc.zero.prop=0.2; out.zero.prop=0.05
var.samp.frac="large"

test.dat=abn.tab.gen1(n.taxa, n.samp.grp1, n.samp.grp2, low.abn, med.abn, high.abn, prop.diff, 
                      abn.seed, obs.seed, struc.zero.prop, out.zero.prop, var.samp.frac)
obs.abn=test.dat$obs.abn
meta.data=cbind(Sample.ID=paste0("sub", seq(n.samp.grp1+n.samp.grp2)), 
                group=rep(c(1, 2), c(n.samp.grp1, n.samp.grp2)))

# Pre-processing
feature.table=obs.abn; sample.var="Sample.ID"; group.var="group"
zero.cut=0.90; lib.cut=1000; neg.lb=FALSE
pre.process=feature_table_pre_process(feature.table, meta.data, sample.var, 
                                      group.var, zero.cut, lib.cut, neg.lb)
feature.table=pre.process$feature.table
group.name=pre.process$group.name
group.ind=pre.process$group.ind
struc.zero=pre.process$structure.zeros

# Paras for ANCOM-BC
grp.name=group.name; grp.ind=group.ind; adj.method="bonferroni"
tol.EM=1e-5; max.iterNum=100; perNum = 1000; alpha=0.05
out<-try(ANCOM_BC(feature.table, grp.name, grp.ind, struc.zero, adj.method, 
                  tol.EM, max.iterNum, perNum, alpha), silent = T)

# ANCOM.BC: Two-stage bias correction
if (inherits(out, "try-error")){
  ANCOM.BC=rep(NA, n.samp.grp1+n.samp.grp2)
}else{ANCOM.BC=out$d}
sub.keep=match(names(ANCOM.BC), colnames(obs.abn))

# True value
ACTUAL=log(test.dat$mult); ACTUAL=ACTUAL[sub.keep]

countdata=test.dat$obs.abn
zero.threshold=0.90
taxa.info.ind=apply(countdata, 1, function(x) sum(x==0)/ncol(countdata))
countdata=countdata[which(taxa.info.ind<zero.threshold), ]+1L
# CSS: Cumulative-sum scaling
meta.data=data.frame(group=rep(c(1, 2), c(n.samp.grp1, n.samp.grp2)))
rownames(meta.data)=paste0("sub", seq(n.samp.grp1+n.samp.grp2))
feature_table=countdata+1
phenotypeData = Biobase::AnnotatedDataFrame(meta.data)
obj = metagenomeSeq::newMRexperiment(countdata, phenoData=phenotypeData, featureData=NULL)
# Calculating normalization factors
p = metagenomeSeq::cumNormStatFast(obj)
obj = metagenomeSeq::cumNorm(obj, p = p)
CSS = metagenomeSeq::normFactors(obj); CSS=log(CSS)[sub.keep]

# MED: Median normalization
coldata=data.frame(group=as.factor(rep(c(1, 2), c(n.samp.grp1, n.samp.grp2))))
rownames(coldata)=paste0("sub", seq(n.samp.grp1+n.samp.grp2))
feature_table=countdata+1
count.table=DESeq2::DESeqDataSetFromMatrix(countData = feature_table, 
                                           colData = coldata, design = ~ group)
dds<-try(DESeq2::DESeq(count.table, quiet = T), silent = T)
if (inherits(dds, "try-error")){
  MED=rep(NA, n.samp.grp1+n.samp.grp2)[sub.keep]
}else{
  MED=DESeq2::sizeFactors(dds)
  MED=log(MED)[sub.keep]}

# UQ: Upper quartile normalization
groupdata=factor(rep(c(1, 2), c(n.samp.grp1, n.samp.grp2)))
feature_table=countdata+1
dds=edgeR::DGEList(counts = feature_table, group = groupdata)
dds=edgeR::calcNormFactors(dds, method="upperquartile")
UQ=dds$samples$norm.factors; UQ=log(UQ)[sub.keep]

# TMM: Trimed mean of m-values
dds=edgeR::DGEList(counts = feature_table, group = groupdata)
dds=edgeR::calcNormFactors(dds, method="TMM")
TMM=dds$samples$norm.factors; TMM=log(TMM)[sub.keep]

# TSS: Total-sum scaling
TSS=colSums(countdata, na.rm = T); TSS=log(TSS)[sub.keep]

norm.df=data.frame(ACTUAL, ANCOM.BC, CSS, MED, UQ, TMM, TSS,
                   group=rep(c("Group1", "Group2"), sapply(group.ind, length)))
norm.df=norm.df%>%mutate(D.ANCOM.BC = ACTUAL - ANCOM.BC,
                         D.CSS = ACTUAL - CSS, D.MED = ACTUAL - MED,
                         D.UQ = ACTUAL - UQ, D.TMM = ACTUAL - TMM, D.TSS = ACTUAL - TSS,
                         CD.ANCOM.BC = scale(D.ANCOM.BC, scale = F),
                         CD.CSS = scale(D.CSS, scale = F), CD.MED = scale(D.MED, scale = F),
                         CD.UQ = scale(D.UQ, scale = F), CD.TMM = scale(D.TMM, scale = F),
                         CD.TSS = scale(D.TSS, scale = F))
write.csv(norm.df, file = "../data/sim_norm/norm_large.csv", row.names = F)
```

## 2.2 Fig. 3a

```{r, message=FALSE, warning=FALSE, comment=NA}
norm.df = read_csv("../data/sim_norm/norm_large.csv")
p.df = norm.df%>%gather(key = "method", value = "value", CD.ANCOM.BC:CD.TSS)
p.df$method = factor(p.df$method, 
                     levels = c("CD.ANCOM.BC", "CD.CSS", "CD.MED", "CD.UQ", "CD.TMM", "CD.TSS"))
p=ggplot(p.df, aes(x=method, y=value, color=method))+ 
  geom_boxplot()+geom_hline(yintercept = 0, linetype="dotted")+
  geom_jitter(color="gray28", position=position_jitter(0.2), aes(shape=group))+
  scale_color_manual(name=NULL,
                     label=c("ANCOM-BC", "CSS", "MED", "UQ", "TMM", "TSS"),
                     values = gg_color_hue(6))+
  scale_shape_manual(name=NULL, values=c(1, 17))+
  labs(x="", y="Centered Deviance")+
  theme_bw()+
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.ticks.x=element_blank(),
        axis.text.x = element_blank(),
        legend.position = c(0.25, 0.85),
        legend.direction = "horizontal", legend.box = "vertical")
ggarrange(p, labels = "a")
ggsave("../figures/Figure 3a.pdf", width=6.25, height=5, units='in')
ggsave("../figures/Figure 3a.jpeg", width=6.25, height=5, units='in', dpi = 300)
```

# 3. Normalization Efficacy with the Presence of Moderate Variable Sampling Fractions

## 3.1 Run a single simulation dataset

```{r, message=FALSE, warning=FALSE, comment=NA, eval=FALSE}
n.taxa=500; n.samp.grp1=30; n.samp.grp2=30; low.abn=50; med.abn=200; high.abn=10000
prop.diff=0.25; abn.seed=123; obs.seed=124; struc.zero.prop=0.2; out.zero.prop=0.05
var.samp.frac="moderate"

test.dat=abn.tab.gen1(n.taxa, n.samp.grp1, n.samp.grp2, low.abn, med.abn, high.abn, prop.diff, 
                      abn.seed, obs.seed, struc.zero.prop, out.zero.prop, var.samp.frac)
obs.abn=test.dat$obs.abn
meta.data=cbind(Sample.ID=paste0("sub", seq(n.samp.grp1+n.samp.grp2)), 
                group=rep(c(1, 2), c(n.samp.grp1, n.samp.grp2)))

# Pre-processing
feature.table=obs.abn; sample.var="Sample.ID"; group.var="group"
zero.cut=0.90; lib.cut=1000; neg.lb=FALSE
pre.process=feature_table_pre_process(feature.table, meta.data, sample.var, 
                                      group.var, zero.cut, lib.cut, neg.lb)
feature.table=pre.process$feature.table
group.name=pre.process$group.name
group.ind=pre.process$group.ind
struc.zero=pre.process$structure.zeros

# Paras for ANCOM-BC
grp.name=group.name; grp.ind=group.ind; adj.method="bonferroni"
tol.EM=1e-5; max.iterNum=100; perNum = 1000; alpha=0.05
out<-try(ANCOM_BC(feature.table, grp.name, grp.ind, struc.zero, adj.method, 
                  tol.EM, max.iterNum, perNum, alpha), silent = T)

# ANCOM.BC: Two-stage bias correction
if (inherits(out, "try-error")){
  ANCOM.BC=rep(NA, n.samp.grp1+n.samp.grp2)
}else{ANCOM.BC=out$d}
sub.keep=match(names(ANCOM.BC), colnames(obs.abn))

# True value
ACTUAL=log(test.dat$mult); ACTUAL=ACTUAL[sub.keep]

countdata=test.dat$obs.abn
zero.threshold=0.90
taxa.info.ind=apply(countdata, 1, function(x) sum(x==0)/ncol(countdata))
countdata=countdata[which(taxa.info.ind<zero.threshold), ]+1L
# CSS: Cumulative-sum scaling
meta.data=data.frame(group=rep(c(1, 2), c(n.samp.grp1, n.samp.grp2)))
rownames(meta.data)=paste0("sub", seq(n.samp.grp1+n.samp.grp2))
feature_table=countdata+1
phenotypeData = Biobase::AnnotatedDataFrame(meta.data)
obj = metagenomeSeq::newMRexperiment(countdata, phenoData=phenotypeData, featureData=NULL)
# Calculating normalization factors
p = metagenomeSeq::cumNormStatFast(obj)
obj = metagenomeSeq::cumNorm(obj, p = p)
CSS = metagenomeSeq::normFactors(obj); CSS=log(CSS)[sub.keep]

# MED: Median normalization
coldata=data.frame(group=as.factor(rep(c(1, 2), c(n.samp.grp1, n.samp.grp2))))
rownames(coldata)=paste0("sub", seq(n.samp.grp1+n.samp.grp2))
feature_table=countdata+1
count.table=DESeq2::DESeqDataSetFromMatrix(countData = feature_table, 
                                           colData = coldata, design = ~ group)
dds<-try(DESeq2::DESeq(count.table, quiet = T), silent = T)
if (inherits(dds, "try-error")){
  MED=rep(NA, n.samp.grp1+n.samp.grp2)[sub.keep]
}else{
  MED=DESeq2::sizeFactors(dds)
  MED=log(MED)[sub.keep]}

# UQ: Upper quartile normalization
groupdata=factor(rep(c(1, 2), c(n.samp.grp1, n.samp.grp2)))
feature_table=countdata+1
dds=edgeR::DGEList(counts = feature_table, group = groupdata)
dds=edgeR::calcNormFactors(dds, method="upperquartile")
UQ=dds$samples$norm.factors; UQ=log(UQ)[sub.keep]

# TMM: Trimed mean of m-values
dds=edgeR::DGEList(counts = feature_table, group = groupdata)
dds=edgeR::calcNormFactors(dds, method="TMM")
TMM=dds$samples$norm.factors; TMM=log(TMM)[sub.keep]

# TSS: Total-sum scaling
TSS=colSums(countdata, na.rm = T); TSS=log(TSS)[sub.keep]

norm.df=data.frame(ACTUAL, ANCOM.BC, CSS, MED, UQ, TMM, TSS,
                   group=rep(c("Group1", "Group2"), sapply(group.ind, length)))
norm.df=norm.df%>%mutate(D.ANCOM.BC = ACTUAL - ANCOM.BC,
                         D.CSS = ACTUAL - CSS, D.MED = ACTUAL - MED,
                         D.UQ = ACTUAL - UQ, D.TMM = ACTUAL - TMM, D.TSS = ACTUAL - TSS,
                         CD.ANCOM.BC = scale(D.ANCOM.BC, scale = F),
                         CD.CSS = scale(D.CSS, scale = F), CD.MED = scale(D.MED, scale = F),
                         CD.UQ = scale(D.UQ, scale = F), CD.TMM = scale(D.TMM, scale = F),
                         CD.TSS = scale(D.TSS, scale = F))
write.csv(norm.df, file = "../data/sim_norm/norm_moderate.csv")
```

## 3.2 Fig. S1a

```{r, message=FALSE, warning=FALSE, comment=NA}
norm.df = read_csv("../data/sim_norm/norm_moderate.csv")
p.df = norm.df%>%gather(key = "method", value = "value", CD.ANCOM.BC:CD.TSS)
p.df$method = factor(p.df$method,
                     levels = c("CD.ANCOM.BC", "CD.CSS", "CD.MED", "CD.UQ", "CD.TMM", "CD.TSS"))
p=ggplot(p.df, aes(x=method, y=value, color=method))+ 
  geom_boxplot()+geom_hline(yintercept = 0, linetype="dotted")+
  geom_jitter(color="gray28", position=position_jitter(0.2), aes(shape=group))+
  scale_color_manual(name=NULL,
                     label=c("ANCOM-BC", "CSS", "MED", "UQ", "TMM", "TSS"),
                     values = gg_color_hue(6))+
  scale_shape_manual(name=NULL, values=c(1, 17))+
  labs(x="", y="Centered Difference")+
  theme_bw()+
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.ticks.x=element_blank(),
        axis.text.x = element_blank(),
        legend.position = c(0.25, 0.85),
        legend.direction = "horizontal", legend.box = "vertical")
ggarrange(p, labels = "a")
ggsave("../figures/Figure S1a.pdf", width=6.25, height=5, units='in')
ggsave("../figures/Figure S1a.jpeg", width=6.25, height=5, units='in', dpi = 300)
```

# 4. Normalization Efficacy with the Presence of Small Variable Sampling Fractions

## 4.1 Run a single simulation dataset

```{r, message=FALSE, warning=FALSE, comment=NA, eval=FALSE}
n.taxa=500; n.samp.grp1=30; n.samp.grp2=30; low.abn=50; med.abn=200; high.abn=10000
prop.diff=0.25; abn.seed=123; obs.seed=124; struc.zero.prop=0.2; out.zero.prop=0.05
var.samp.frac="small"

test.dat=abn.tab.gen1(n.taxa, n.samp.grp1, n.samp.grp2, low.abn, med.abn, high.abn, prop.diff, 
                      abn.seed, obs.seed, struc.zero.prop, out.zero.prop, var.samp.frac)
obs.abn=test.dat$obs.abn
meta.data=cbind(Sample.ID=paste0("sub", seq(n.samp.grp1+n.samp.grp2)), 
                group=rep(c(1, 2), c(n.samp.grp1, n.samp.grp2)))

# Pre-processing
feature.table=obs.abn; sample.var="Sample.ID"; group.var="group"
zero.cut=0.90; lib.cut=1000; neg.lb=FALSE
pre.process=feature_table_pre_process(feature.table, meta.data, sample.var, 
                                      group.var, zero.cut, lib.cut, neg.lb)
feature.table=pre.process$feature.table
group.name=pre.process$group.name
group.ind=pre.process$group.ind
struc.zero=pre.process$structure.zeros

# Paras for ANCOM-BC
grp.name=group.name; grp.ind=group.ind; adj.method="bonferroni"
tol.EM=1e-5; max.iterNum=100; perNum = 1000; alpha=0.05
out<-try(ANCOM_BC(feature.table, grp.name, grp.ind, struc.zero, adj.method, 
                  tol.EM, max.iterNum, perNum, alpha), silent = T)

# ANCOM.BC: Two-stage bias correction
if (inherits(out, "try-error")){
  ANCOM.BC=rep(NA, n.samp.grp1+n.samp.grp2)
}else{ANCOM.BC=out$d}
sub.keep=match(names(ANCOM.BC), colnames(obs.abn))

# True value
ACTUAL=log(test.dat$mult); ACTUAL=ACTUAL[sub.keep]

countdata=test.dat$obs.abn
zero.threshold=0.90
taxa.info.ind=apply(countdata, 1, function(x) sum(x==0)/ncol(countdata))
countdata=countdata[which(taxa.info.ind<zero.threshold), ]+1L
# CSS: Cumulative-sum scaling
meta.data=data.frame(group=rep(c(1, 2), c(n.samp.grp1, n.samp.grp2)))
rownames(meta.data)=paste0("sub", seq(n.samp.grp1+n.samp.grp2))
feature_table=countdata+1
phenotypeData = Biobase::AnnotatedDataFrame(meta.data)
obj = metagenomeSeq::newMRexperiment(countdata, phenoData=phenotypeData, featureData=NULL)
# Calculating normalization factors
p = metagenomeSeq::cumNormStatFast(obj)
obj = metagenomeSeq::cumNorm(obj, p = p)
CSS = metagenomeSeq::normFactors(obj); CSS=log(CSS)[sub.keep]

# MED: Median normalization
coldata=data.frame(group=as.factor(rep(c(1, 2), c(n.samp.grp1, n.samp.grp2))))
rownames(coldata)=paste0("sub", seq(n.samp.grp1+n.samp.grp2))
feature_table=countdata+1
count.table=DESeq2::DESeqDataSetFromMatrix(countData = feature_table, 
                                           colData = coldata, design = ~ group)
dds<-try(DESeq2::DESeq(count.table, quiet = T), silent = T)
if (inherits(dds, "try-error")){
  MED=rep(NA, n.samp.grp1+n.samp.grp2)[sub.keep]
}else{
  MED=DESeq2::sizeFactors(dds)
  MED=log(MED)[sub.keep]}

# UQ: Upper quartile normalization
groupdata=factor(rep(c(1, 2), c(n.samp.grp1, n.samp.grp2)))
feature_table=countdata+1
dds=edgeR::DGEList(counts = feature_table, group = groupdata)
dds=edgeR::calcNormFactors(dds, method="upperquartile")
UQ=dds$samples$norm.factors; UQ=log(UQ)[sub.keep]

# TMM: Trimed mean of m-values
dds=edgeR::DGEList(counts = feature_table, group = groupdata)
dds=edgeR::calcNormFactors(dds, method="TMM")
TMM=dds$samples$norm.factors; TMM=log(TMM)[sub.keep]

# TSS: Total-sum scaling
TSS=colSums(countdata, na.rm = T); TSS=log(TSS)[sub.keep]

norm.df=data.frame(ACTUAL, ANCOM.BC, CSS, MED, UQ, TMM, TSS,
                   group=rep(c("Group1", "Group2"), sapply(group.ind, length)))
norm.df=norm.df%>%mutate(D.ANCOM.BC = ACTUAL - ANCOM.BC,
                         D.CSS = ACTUAL - CSS, D.MED = ACTUAL - MED,
                         D.UQ = ACTUAL - UQ, D.TMM = ACTUAL - TMM, D.TSS = ACTUAL - TSS,
                         CD.ANCOM.BC = scale(D.ANCOM.BC, scale = F),
                         CD.CSS = scale(D.CSS, scale = F), CD.MED = scale(D.MED, scale = F),
                         CD.UQ = scale(D.UQ, scale = F), CD.TMM = scale(D.TMM, scale = F),
                         CD.TSS = scale(D.TSS, scale = F))
write.csv(norm.df, file = "../data/sim_norm/norm_small.csv")
```

## 4.2 Fig. S2a

```{r, message=FALSE, warning=FALSE, comment=NA}
norm.df = read_csv("../data/sim_norm/norm_small.csv")
p.df = norm.df%>%gather(key = "method", value = "value", CD.ANCOM.BC:CD.TSS)
p.df$method = factor(p.df$method, 
                     levels = c("CD.ANCOM.BC", "CD.CSS", "CD.MED", "CD.UQ", "CD.TMM", "CD.TSS"))
p=ggplot(p.df, aes(x=method, y=value, color=method))+ 
  geom_boxplot()+geom_hline(yintercept = 0, linetype="dotted")+
  geom_jitter(color="gray28", position=position_jitter(0.2), aes(shape=group))+
  scale_color_manual(name=NULL,
                     label=c("ANCOM-BC", "CSS", "MED", "UQ", "TMM", "TSS"),
                     values = gg_color_hue(6))+
  scale_shape_manual(name=NULL, values=c(1, 17))+
  labs(x="", y="Centered Difference")+
  theme_bw()+
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.ticks.x=element_blank(),
        axis.text.x = element_blank(),
        legend.position = c(0.25, 0.85),
        legend.direction = "horizontal", legend.box = "vertical")
ggarrange(p, labels = "a")
ggsave("../figures/Figure S2a.pdf", width=6.25, height=5, units='in')
ggsave("../figures/Figure S2a.jpeg", width=6.25, height=5, units='in', dpi = 300)
```

# Session information

```{r, message=FALSE, warning=FALSE, comment=NA}
sessionInfo()
devtools::session_info()
```







